---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mcgf

<!-- badges: start -->
<!-- badges: end -->

The goal of `mcgf` is to provide easy-to-use functions for simulating and fitting covariance models. It provides functions for simulating (regime-switching)  Markov chain Gaussian fields with covariance functions of the Gneiting class by simple kriging. Parameter estimation methods such as weighted least squares and maximum likelihood estimation are available.

## Installation

You can install the development version of mcgf from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("tianxia-jia/mcgf")
```

## Data Simulation

To simulate a MCGF with fully symmetric covariance structure, we begin with simulating 10 locations randomly.

```{r}
library(mcgf)
set.seed(1234)
dists <- rdists(10)
```

Next, we simulate a MCGF with the general stationary covariance structure. In this example the covariance structure is a convex combination of a base separable model and a Lagrangian model account for asymmetry.

```{r, message = F, warning=F}
library(mcgf)
par_spatial <- list(nugget = 0, c = 0.01, gamma = 0.3)
par_temporal <- list(a = 1, alpha = 0.5)
par_sep <- list(par_s = par_spatial, par_t = par_temporal)
par_lagr <- list(v1 = 10, v2 = 50, k = 2)

set.seed(1234)
X <- mcgf_sim(N = 1500, base = "sep", lagrangian = "lagr_tri", lambda = 0.5, 
              par_base = par_sep, par_lagr = par_lagr, dists = dists, 
              lag = 10)
plot.ts(X)
```

## Parameter Estimation

### Create an `mcgf` object
To estimate parameters, we need to calculate auto-correlations and cross-correlations. Let's first create an `mcgf` object. The `mcgf` class extends the `data.frame` with more attributes.

```{r}
train <- X[1:1000, ]
test <- X[-c(1:1000), ]
x_mcgf <- mcgf(train, dists = dists)
```

Then the acfs and ccfs can be added to this object as follows.

```{r}
x_mcgf <- add_acfs(x = x_mcgf, lag_max = 10)
x_mcgf <- add_ccfs(x = x_mcgf, lag_max = 10)
```

### Estimate base model

To perform parameter estimation, we can start with estimating the parameters for spatial and temporal models.

```{r}
fit_spatial <- fit_base(x = x_mcgf, lag = 10, model = "spatial", 
                        par_init = list(nugget = 0, c = 0.001, gamma = 0.5), 
                        method = "wls")
fit_spatial$fit
```
```{r}
fit_temporal <- fit_base(x = x_mcgf, lag = 10, model = "temporal", 
                        par_init = list(a = 1, alpha = 0.5), 
                        method = "wls")
fit_temporal$fit
```
Alternatively, we can fit the separable model all at once:

```{r}
fit_sep <- fit_base(x = x_mcgf, lag = 10, model = "sep",
                    par_init = list(a = 1, alpha = 0.5, nugget = 0, c = 0.001, 
                         gamma = 0.5))
fit_sep$fit
```
we can also estimate the parameters using MLE:
```{r}
fit_sep2 <- fit_base(x = x_mcgf, lag = 10, model = "sep",
                    par_init = list(a = 1, alpha = 0.5, nugget = 0, c = 0.001, 
                         gamma = 0.5), method = "mle")
fit_sep2$fit
```

Now we will add the base model to `x_mcgf`:
```{r}
x_mcgf <- add_base(x_mcgf, fit_base = fit_sep)
```

To print the current model, we do
```{r}
model(x_mcgf)
```

### Estimate the Lagrangian model

Similarly, we can estimate the parameters for the Lagrangian component by
```{r}
fit_lagr <- fit_lagr(x = x_mcgf, model = "lagr_tri",
                    par_init = list(lambda = 0.1, v1 = 10, v2 = 10, k = 1))
fit_lagr$fit
```

We can add the Lagrangian model by
```{r}
x_mcgf <- add_lagr(x_mcgf, fit_lagr = fit_lagr)
```

Finally we may print the final model:
```{r}
model(x_mcgf)
```

### Kriging forecast

This package provides kriging forecasts (and intervals) for empirical, base, and general stationary models. We will obtain forecasts for the test dataset.
```{r}
# Empirical model
fit_emp <- krige(x_mcgf, model = "empirical", newdata = test, interval = TRUE)
rmse_emp <- sqrt(mean(colMeans((test - fit_emp$fit) ^ 2, na.rm = T)))

# Base separable model
fit_base <- krige(x_mcgf, model = "base", newdata = test, interval = TRUE)
rmse_base <- sqrt(mean(colMeans((test - fit_base$fit) ^ 2, na.rm = T)))

# Stationary model
fit_stat <- krige(x_mcgf, model = "all", newdata = test, interval = TRUE)
rmse_stat <- sqrt(mean(colMeans((test - fit_stat$fit) ^ 2, na.rm = T)))

rmse <- c(rmse_emp, rmse_base, rmse_stat)
names(rmse) <- c("Empirical", "Separable", "Stationary")
rmse
```

