---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mcgf

<!-- badges: start -->
<!-- badges: end -->

The goal of `mcgf` is to provide easy-to-use functions for simulating and fitting covariance models. This package has functions to simulate samples for Markov chain Gaussian fields with covariance functions of the Gneiting class by simple kriging. Parameter estimation methods such as weighted least squares and maximum likelihood estimation are available.

## Installation

You can install the development version of mcgf from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("tianxia-jia/mcgf")
```

## Data Simulation

To simulate a MCGF with fully symmetric covariance structure, we first create distance arrays for 10 locations:

```{r}
library(mcgf)
set.seed(123)
dists <- rdists(10)
```

Next, we simulate a MCGF.

```{r}
library(mcgf)
par_s <- list(nugget = 0.2, c = 0.001, gamma = 0.3)
par_t <- list(a = 5, alpha = 0.5)
par_base <- list(par_s = par_s, par_t = par_t)

set.seed(123)
X <- mcgf_sim(N = 1000, base = "sep", lagrangian = "none", par_base = par_base,
              dists = dists, lag = 10)
plot.ts(X)
```

## Parameter Estimation

To estimate parameters using weighted least squares method, we need to calculate autocorrelations and cross-autocorrelations. Let's first create a `mcgf` object. The `mcgf` class extends the `data.frame` with more attributes.

```{r}
x_mcgf <- mcgf(X, dists = dists)
```

Then the acfs and ccfs can be added to this object as follows.

```{r}
x_mcgf <- add_acfs(x = x_mcgf, lag_max = 10)
x_mcgf <- add_ccfs(x = x_mcgf, lag_max = 10)
```

To find the weighted least squares, we can run

```{r}
fit_base(x = x_mcgf, lag = 10, model = "sep", optim_fn = "nlminb",
         par_init = list(a = 1, alpha = 0.5, nugget = 0.1, c = 0.001, 
                         gamma = 0.5), method = "wls")
```
To find the maximum likelihood estimates, we run

```{r}
fit_base(x = x_mcgf, lag = 10, model = "sep", optim_fn = "nlminb",
         par_init = list(a = 1, alpha = 0.5, nugget = 0.1, c = 0.001, 
                         gamma = 0.5), method = "mle")
```

<!-- ## Regime-Switching MCGF -->

<!-- Regime-switching MCGF can be simulated as -->

<!-- ```{r, warning=FALSE,error=FALSE} -->
<!-- par_s <- list(nugget = 0.5, c = 0.01, gamma = 0.5) -->
<!-- par_t <- list(a = 1, alpha = 0.5) -->
<!-- par_base <- list(par_s = par_s, par_t = par_t) -->
<!-- par_lagr <- list(v1 = 5, v2 = 10) -->

<!-- set.seed(123) -->
<!-- labels <- sample(1:2, 1000, replace = TRUE) -->
<!-- X <- mcgf_rs_sim(N = 1000, -->
<!--                  labels = labels, -->
<!--                  base_ls = list("sep"), -->
<!--                  lagrangian_ls = list("none", "none"), -->
<!--                  lambda_ls = list(0, 0.5), -->
<!--                  par_base_ls = list(par_base), -->
<!--                  par_lagr_ls = list(NULL, par_lagr), -->
<!--                  dists_ls = list(dists, dists)) -->
<!-- plot.ts(X[, -1]) -->
<!-- ``` -->

