---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# mcgf

<!-- badges: start -->
<!-- badges: end -->

The goal of `mcgf` is to provide easy-to-use functions for simulating and fitting covariance models. It provides functions for simulating (regime-switching)  Markov chain Gaussian fields with covariance functions of the Gneiting class by simple kriging. Parameter estimation methods such as weighted least squares and maximum likelihood estimation are available.

## Installation

You can install the development version of mcgf from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("tianxia-jia/mcgf")
```


## Data Simulation

To simulate a MCGF with fully symmetric covariance structure, we begin with simulating 10 locations randomly.

```{r, message = F, warning=F}
library(mcgf)
set.seed(123)
h <- rdists(10)
```

Next, we simulate a MCGF with the general stationary covariance structure. In this example the covariance structure is a convex combination of a base separable model and a Lagrangian model account for asymmetry.

```{r, message = F, warning=F}
N <- 1000
lag <- 5

par_base <- list(
    par_s = list(nugget = 0, c = 0.001, gamma = 0.5),
    par_t = list(a = 0.5, alpha = 0.8)
)
par_lagr <- list(v1 = 200, v2 = 200, k = 2)

sim1 <- mcgf_sim(
    N = N,
    base = "sep",
    lagrangian = "lagr_tri",
    par_base = par_base,
    par_lagr = par_lagr,
    lambda = 0.2,
    dists = h,
    lag = lag
)
sim1 <- sim1[-c(1:(lag + 1)), ]
rownames(sim1) <- 1:nrow(sim1)

sim1 <- list(data = sim1, dists = h)
```

## Parameter Estimation

### Create an `mcgf` object
To estimate parameters, we need to calculate auto-correlations and cross-correlations. Let's first create an `mcgf` object. The `mcgf` class extends the `data.frame` with more attributes.
```{r}
sim1_mcgf <- mcgf(sim1$data, dists = sim1$dists)
```

Then the acfs and ccfs can be added to this object as follows.

```{r}
sim1_mcgf <- add_acfs(sim1_mcgf, lag_max = lag)
sim1_mcgf <- add_ccfs(sim1_mcgf, lag_max = lag)
```

### Estimate base model

To perform parameter estimation, we can start with estimating the parameters for spatial and temporal models.

```{r}
fit_spatial <- fit_base(
    sim1_mcgf,
    model = "spatial",
    lag = lag,
    par_init = c(c = 0.001, gamma = 0.5),
    par_fixed = c(nugget = 0)
)
fit_spatial$fit
```
```{r}
fit_temporal <- fit_base(
    sim1_mcgf,
    model = "temporal",
    lag = lag,
    par_init = c(a = 0.3, alpha = 0.5)
)
fit_temporal$fit
```
Alternatively, we can fit the separable model all at once:

```{r}
fit_sep <- fit_base(
    sim1_mcgf,
    model = "sep",
    lag = lag,
    par_init = c(
        c = 0.001,
        gamma = 0.5,
        a = 0.5,
        alpha = 0.5
    ),
    par_fixed = c(nugget = 0)
)
fit_sep$fit
```
we can also estimate the parameters using MLE:
```{r}
fit_sep2 <- fit_base(
    sim1_mcgf,
    model = "sep",
    lag = lag,
    par_init = c(
        c = 0.001,
        gamma = 0.5,
        a = 0.5,
        alpha = 0.5
    ),
    par_fixed = c(nugget = 0),
    method = "mle",
)
fit_sep2$fit
```

Now we will add the base model to `x_mcgf`:
```{r}
sim1_mcgf <- add_base(sim1_mcgf, fit_base = fit_sep)
```

To print the current model, we do
```{r}
model(sim1_mcgf)
```

### Estimate the Lagrangian model

Similarly, we can estimate the parameters for the Lagrangian component by
```{r}
fit_lagr <- fit_lagr(
    sim1_mcgf,
    model = "lagr_tri",
    par_init = c(v1 = 300, v2 = 300, lambda = 0.15),
    par_fixed = c(k = 2)
)
fit_lagr$fit
```

We can add the Lagrangian model by
```{r}
sim1_mcgf <- add_lagr(sim1_mcgf, fit_lagr = fit_lagr)
```

Finally we may print the final model:
```{r}
model(sim1_mcgf)
```

### Kriging forecast

This package provides kriging forecasts (and intervals) for empirical, base, and general stationary models.
```{r}
# Empirical model
fit_emp <-
    krige(sim1_mcgf,
        model = "empirical",
        interval = TRUE
    )
rmse_emp <- sqrt(mean(colMeans((sim1_mcgf - fit_emp$fit)^2, na.rm = T)))

# Base separable model
fit_base <-
    krige(sim1_mcgf,
        model = "base",
        interval = TRUE
    )
rmse_base <-
    sqrt(mean(colMeans((sim1_mcgf - fit_base$fit)^2, na.rm = T)))

# Stationary model
fit_stat <-
    krige(sim1_mcgf,
        model = "all",
        interval = TRUE
    )
rmse_stat <-
    sqrt(mean(colMeans((sim1_mcgf - fit_stat$fit)^2, na.rm = T)))

rmse <- c(rmse_emp, rmse_base, rmse_stat)
names(rmse) <- c("Empirical", "Separable", "Stationary")
rmse
```

